---
title: "WUP_04"
author: "Jaci Heckman"
date: "Summer 2024"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE,echo=FALSE}
require(knitr)
require(tidyverse)
require(tigerstats)
mydata<- read.csv(file= "GSScsv.csv") 
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data is in the file GSScsv.csv and came from the 2018 General Social Survey.  The question to answer is to resolve this hypothesis.  
$H_0$ There is no difference between the average score on the WORDSUM test and level of schooling DEGREE  
$H_a$ There is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE  

# Methods

<!--Mention that you intend to use chatGPT to the greatest extent possible. Give credit to chatGPT when you use it to create code in your report. Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
boxplot(WORDSUM~DEGREE,data=mydata)
```
  
In this preliminary display of the three data formats, Bachelor, Graduate, and Junior College levels. The initial inference would be that there is a distinct height difference in the levels of academic achievements. The lowest levels are in the junior college sections. We can determine that this is so by looking at black line in the boxes, the median, and we can determine that there is only few instances of overlap per each pair of data boxes. 
  
```{r}
basicCN <- ggplot(mydata,aes(x=DEGREE,y=WORDSUM))
```
```{r}
basicCN + geom_boxplot()
```
    
These two data displays show multiple boxes placed at levels from 2 through 13, with five sets of academic levels listed on the bottom row. This display shows that the LT high school level, along with the high school box shared the same bottom line at approximately, 7.2. There is indeed a difference in the height of the five data points. The second pictograph showcases this element best, the same information was included and the same data was used to reach all ten box placements.  

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->
  
```{r}
favstats(WORDSUM~DEGREE,data=mydata)
```
  
In this sample excerpt of information for the five data sets, there is a displayed quadrant, using the letter Q and a number, 1 or 3. This number quantifies the itemized information into quarters which shows the percentile or whole number that his portion of data contains. This information, when you compare the largest Q1 number and the Q3 reports, you can see a large deviance in information. Specifically, the LT High school Q1 report, 7 compared to the Bachelor Q3 number 11. To assume that this means there is a statistical element, there will need to be an inferential test. To complete this, we will utilize tukey HDS. Using this model we can compare pairs of data to each other and can see the information organized much better. 

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
model1 <- aov(WORDSUM~DEGREE,data=mydata)
summary.aov(model1)
```
      
This model information chart through the Anova program showcases the data and proposes the answer to this hypothesis. The probability stats are shown under the Pr item. The probability level would need to exceed 0.5 percent to nullify the hypothesis. In this instance we have a number 0.000236 probability. 


```{r}
#we specify the margins starting at the bottom and proceeding clockwise - thus 8.7 is the left side margin to hold the names of the factors
par(mar = c(3, 8.7, 2, 2))
plot(TukeyHSD(model1),las=2)
``` 
  
This probability display is too low to accept, we reject the null hypothesis and are in favor of the results here. The brackets each exhibit two a pair of the five items and compare them. The top most bracket for example is the Graduate compared to the Bachelor data. We can determine whether or not there is a negative or positive inference by noting where the brackets are placed, on the dotted line (0) or on the left (positive) side or the right (negative) side. There are seven brackets that include at least one comparison sample in the negative zone. This shows that there are very large data differences by looking at the brackets on the very far left side. 

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
  
To conclude the above results, the comparison of data responds positively to the hypothesis. This conclusion was obtained by analyzing numerical reports and graphical data, using plotted box displays and brackets. The ggplot system was used to showcase the level of education on a single plain graph and then again by a multiple system test through the Anova program. Afterwards the quadrant and probability numbers were included and evaluted to show the comparison to the data. There is a positive inference between some levels of higher education in tandem with the WORDSUM test. This includes the Bachelor, Graduate and Junior College academic levels.  
